### 2008-08-29 htdocs/sections.tex
Added porous media information
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig/trunk@406 40876441-3496-0410-a613-a2fca35b2ae2

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  sections.tex
%
%  Content for Ludwig documentation.
%
%  $Id: sections.tex,v 1.2 2008-02-25 09:53:37 kevin Exp $
%
%  Edinburgh Soft Matter and Statistical Physics Group and
%  Edinburgh Parallel Computing Centre
%
%  Kevin Stratford (kevin@epcc.ed.ac.uk)
%  (c) 2008 The University of Edinburgh
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{Ongoing Development}

\subsubsection{Some notes}

\begin{enumerate}

\item
Serial timer is exhausted at about 35 minutes.

\item
Size of the colloid message buffers needs to be computed robustly
at run time, or, allocated on a just-in-time basis with knowledge
of exact size required. The former would mean knowing the number
of particles before the call to CCOM\_init\_halo.

\subsubsection{Other Agenda Items}
\begin{enumerate}

\item
The particle code needs a critic from someone other than myself.
Again we need to abstract it away a bit more.

\item
Two-dimensional model (D2Q9).

\item
Finite Difference for order parameter.

\item
I/O via XML/HDF5.

@@

@@

@@

@@

\bibitem{lv01}
Ladd, A.J.C., and R. Verberg,
Lattice-Boltzmann simulations of particle-fluid suspensions,
\textit{J. Stat. Phys.}, \textbf{104}, 1191, 2001

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  sections.tex
%
%  Content for Ludwig documentation.
%
%  $Id: sections.tex,v 1.3 2008-08-29 09:28:08 kevin Exp $
%
%  Edinburgh Soft Matter and Statistical Physics Group and
%  Edinburgh Parallel Computing Centre
%
%  Kevin Stratford (kevin@epcc.ed.ac.uk)
%  (c) 2008 The University of Edinburgh
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{Download}

@@

\subsection{Porous Media}

Porous media calculations can be undertaken when appropriate
solid/fluid status information is supplied. The are a number
of switches available in the input file:

\inputkey{porous\_media\_file}

specifies the file stub name to be read at the start of execution.
(If the stub name is \texttt{file} then the code will expect to
find \texttt{file.001-001} in the current directory.

\inputkey{porous\_media\_format}

is either \texttt{ASCII} or \texttt{BINARY} as appropriate. Note that
in parallel, a single data file can be supplied, but it must be binary.
The default is \texttt{BINARY}.

\inputkey{porous\_media\_type}

is either \texttt{status\_only} (the default) or \texttt{status\_with\_h}.

\subsubsection{File format}

In all cases the status (fluid/solid) information is represented by a
single \texttt{char} (or integer), which must be supplied via the
porous media file. A single file should contain data matching the
current system size, and have the $z-$direction running fastest,
followed by the $y-$direction. Note that in non-periodic directions,
the structure must be 'closed', i.e., all the points at the edge
should be solid.

Fluid sites are designated by \texttt{0} and boundary or solid sites
by \texttt{1}. These data should be of type \texttt{char} in binary,
and may be integer in ASCII.

Where wetting information is required, the free energy parameter $H$
can be supplied by using the \texttt{status\_with\_h} switch. In this
case, the \texttt{char} status is augmented by a single \texttt{double}
value which is the local value of $H$. The order is then
$s_1,h_1, s_2, h_2, \ldots$.

\subsubsection{Permeability calculations}

Single fluid permeability calculations for a given proous structure
can be undertaken by driving a flow via the fluid body force. Note
that the structure must be periodic in the direction of the force
(this may mean duplicating a 'reflected' version of a given sample
to create the correct input). The body force can be specified so
that, e.g., to drive a flow in the $x-$direction

\texttt{force 0.001\_0.000\_0.000}

The force should not be so large that the maximum velocity generated
threatens the Mach number constraint. To get a measurement of the
flow at equilibrium, the calculation should be run at least the
momentum diffusion time for the system ($L^2/\eta$ in LB time steps).

The net flow can be measured by combining the statistics for the
total momemtum and the total density (which is equal to the volume
with $\rho_0 = 1$).

Note that in the case of porous media with narrow channels at the grid
scale, the wall velocity can be dependent on the viscosity of the fluid.
This is an artefact of the bounce-back on links and will result in a
viscosity-dependent permeability (see, e.g., \cite{lipanmiller}). To minimise
this effect, it is recommended to to on the correction to the ghost
mode relaxation time proposed by Ginzburg and d'Humi\`eres \cite{ginzburg}
which corrects the viscosity-dependence of the apparent boundary
position. This is done via the input switches

\texttt{ghost\_modes on\\
ginzburg\_dhumieres on}

and should recover a viscosity-independent premeability measurement
\cite{lipanmiller,chunladd}.

\subsubsection{Capillaries}

The utility \texttt{capillary.c} can be used to generate various
porous media input files corresponding to various simple square
or circular capillaries.

@@

\bibitem{chunladd}
B. Chun and A.J.C. Ladd,
Interpolated boundary condition for lattice Boltzmann simulations in
narrow gaps,
\textit{Phys. Rev. E}, \textbf{75}, 066705, 2007.

@@

\bibitem{ginzburg}
I. Ginzburg and D. d'Humi\`eres,
Multireflection boundary conditions for lattice Boltzmann models,
\textit{Phys. Rev. E}, \textbf{68}, 066614, 2003.

@@

\bibitem{lv01}
Ladd, A.J.C., and R. Verberg,
Lattice-Boltzmann simulations of particle-fluid suspensions,
\textit{J. Stat. Phys.}, \textbf{104}, 1191, 2001.

\bibitem{lipanmiller}
H. Li, C. Pan, and C.T. Miller,
Pore-scale investigation of viscous coupling effects for two-phase
flow in porous media,
\textit{Phys. Rev. E}, \textbf{72}, 026705, 2005.
